Providing a Unified Software Environment for Canada’s National Advanced Computing Centers

### 2019-04-09 paper.tex
Merged in new changes from Ryan and Maxime

The second principle is that the software environment should be decoupled from the underlying operating system. Since each tool for software package installation has strengths and weaknesses, we adopted a layered approach to benefit from the best tool for each kind of software package. We briefly described this in Section~\ref{sec:Layered_environment}.

@@

Finally, we describe in Section~\ref{sec:Making_it_available} how you can make use of our software stack on your computer or cluster, mention other marginal benefits from our design in Section~\ref{sec:Other_benefits}.

@@

One final use case for CVMFS is to distribute datasets. By default, CVMFS  is optimized for a large number of small files, rather than a small number of large files, so different  configurations and access methods are needed for data repositories. At the time of writing this paper, this is used by some of our research groups in genomics, but not directly by our staff.

@@

The framework contains all of the core code which drives the logic of installing a software package. It includes the following steps: parsing the configuration options, downloading source files, patching them, configuring, building and installing the software, as well as performing post-installation steps such as sanity checks, generating a module or running custom commands.

@@

EasyBuild also automatically creates module files~\cite{Modules1991,Modules1996,Lmod}. The framework supports multiple standard and custom naming schemes. We use a version of a hierarchical naming scheme, in which modules are organized in a tree structure, hiding those that are incompatible with currently loaded modules. This limits the list of visible modules which could otherwise be very extensive. This can however confuse some users, since they don’t automatically see all available modules. Lmod~\cite{Lmod} alleviates this issue by providing a separate command to crawl the full module tree and instructs users on how to load the module they want.

@@

\section{Making the software environment available on your system}
\label{sec:Making_it_available}
Making our software environment available on a single computer is as simple as installing the CVMFS client and configuring repositories manually or installing our RPM
\textbf{TODO: Document our RPM}

@@

If you plan to use our CVMFS repositories on a large number of nodes such as a cluster, we require that you deploy one or more caching HTTP proxies, to avoid overloading the stratum 1 servers. If you are interested in testing or making our software environment available to your users on your cluster, we ask that you first contact us by writing to 
\begin{center}
\url{rsnt-software@computecanada.ca}    
\end{center}
Note that this software stack comes with absolutely no warranty.

\section{Other benefits and conclusion}
\label{sec:Conclusion}
We have built a software environment that is portable, is optimized for multiple CPU architectures and network interconnects, and can be deployed in a matter of minutes on a single node, and not much longer for a whole cluster. The design of the infrastructure guarantees that the software environment seen by our end users is the same across all of our clusters, and will not diverge over time. This makes for a better user experience, as it becomes much easier to migrate between two clusters. It also reduces the workload on our staff by unifying the installation process for all clusters, and makes supporting our users easier because all clusters work the same way.

Because of the portability of our software stack and the distribution mechanism, it can be used on any cluster, or any virtual machine on a cloud. For example, for training events, we have been using it to provision small replicas of Compute Canada clusters in short-lived virtual clusters deployed on our OpenStack platform. We have also used this portable software environment to expose a cluster environment directly through a browser, via JupyterHub. It can also help support infrastructures owned by various institutions. This is the case of many universities in Canada that used to host a cluster which is now used by the institution only, or universities that deploy new local infrastructure and don't have the human resources to build a large software stack.

>>>>

The second principle is that the software environment should be decoupled from the underlying operating system. Since each tool for software package installation has strengths and weaknesses, we adopted a layered approach to leverage the best tool for each kind of software package. We briefly describe this in Section~\ref{sec:Layered_environment}.

@@

Finally, we describe in Section~\ref{sec:Making_it_available} how you can use our software stack on your computer or cluster, and summarize the benefits of our design in Section~\ref{sec:Conclusion}.

@@

One additional use case for CVMFS is to distribute datasets. By default, CVMFS  is optimized for a large number of small files, rather than a small number of large files, so different  configurations and access methods are needed for data repositories. At the time of writing this paper, this is used by some of our research groups in genomics, but not directly by our staff.

@@

The framework contains all of the core code which drives the logic of the entire software installation process, including parsing the configuration options, downloading source files, patching them, configuring, building and installing the software, as well as performing post-installation steps such as sanity checks, generating a module or running custom commands.

@@

EasyBuild also automatically creates module files~\cite{Modules1991,Modules1996,Lmod}. The framework supports multiple standard and custom naming schemes. We use a version of a hierarchical naming scheme, in which modules are organized in a tree structure, hiding those that are incompatible with currently loaded modules. This limits the list of visible modules, which could otherwise be very extensive. This can however confuse some users, since they don’t automatically see all available modules. Lmod~\cite{Lmod} alleviates this issue by providing a separate command to crawl the full module tree and instructs users on how to load the module they want.

@@

\section{Making the software environment available on your system}
\label{sec:Making_it_available}
Making our software environment available on a single computer is as simple as installing the CVMFS client and configuring repositories manually or installing our RPM. This is described on this webpage:
\begin{center}
    \url{https://docs.computecanada.ca/wiki/Accessing_our_software_environment_on_your_computer}
\end{center}

@@

If you plan to use our CVMFS repositories on a large number of nodes such as a cluster, we require that you deploy one or more caching HTTP proxies, to avoid overloading the stratum 1 servers. If you are interested in testing or making our software environment available to your users on your cluster, we ask that you first contact us as described in the webpage mentioned before. Note that this software stack comes with absolutely no warranty.

\section{Conclusion}
\label{sec:Conclusion}
We have built a software environment that is portable, is optimized for multiple CPU architectures and network interconnects, and can be deployed in a matter of minutes on a single node, and not much longer for a whole cluster. The design of the infrastructure guarantees that the software environment seen by our end users is the same across all of our clusters, and will not diverge over time. This makes it much easier for users to use multiple clusters or migrate between clusters, improving the user experience. It also reduces the workload on our staff by unifying the installation process for all clusters, and makes supporting our users easier because all clusters work the same way.

Because of the portability of our software stack and the distribution mechanism, it can be used on any cluster, or any Linux computer or virtual machine on a cloud. For example, for training events, we have been using it to provision small replicas of Compute Canada clusters in short-lived virtual clusters deployed on our OpenStack platform. We have also used this portable software environment to expose a cluster environment directly through a browser, via JupyterHub.

Since software installation is no longer tightly coupled to any particular site or system, this approach has reduced the time required to commission major new Compute Canada clusters and make them accessible to our users by several months. At the same time, it also benefits universities and other research organizations running smaller or older clusters, desktops, and laboratory computing infrastructure, which may not have the human resources to build and maintain a comprehensive research software stack.
